A cascaded supervised learning approach to inverse reinforcement learning

### 2013-06-25 piot2012cascading_ECML/Cascading.tex
Corrections ECML pour le Camera Ready

IRL is one of many ways to perform Apprenticeship Learning (AL): imitating a demonstrated expert policy, without necessarily explicitely looking for the reward function. The reward function nevertheless is of interest in its own right. As mentionned in \cite{russell1998learning}, its semantics can be analyzed in biology or econometrics for instance. Practically, the reward can be seen as a succinct description of a task. Discovering it removes the coupling that exists in AL between understanding the task and learning how to fulfill it. IRL allows the use of (A)DP or RL techniques to learn the task from the computed reward function. A very straightforward non-IRL way to do AL is for example to use a multi-class classifier to directly learn the expert policy. We provide in the experiments (Sec.~\ref{sec:experiments}) a comparison between AL and IRL algorithms by using IRL as a way to do AL.

@@

This algorithm does not need to iteratively solve an MDP and requires only sampled transitions from expert and non-expert policies as inputs. Moreover, up to the use of some heuristics (see Sec.~\ref{subsubsec:heuristics}), the algorithm is able to be trained only with transitions sampled from the demonstrated expert policy. A specific instantiation of CSI (proposed Sec.~\ref{subsubsec:subroutines}) is tested on the mountain car problem (Sec.~\ref{subsec:mountaincar}) and on a highway driving simulator (Sec.~\ref{subsec:highway}) where we compare it with a pure classification algorithm \cite{taskar2005learning} and with two recent successfull IRL methods \cite{klein2012scirl} as well as with a random baseline.

Differences and similarities with existing AL or IRL approaches are succinctly discussed in Sec.~\ref{sec:related}.
\section{Background and notations}
\label{sec:background}
First, we introduce some general notations.
Let $E$ and $F$ be two non-empty sets, $E^F$ is the set of functions from $F$ to $E$.
We note $\Delta_X$ the set of distributions over $X$.
Let $\alpha\in\mathbb{R}^X$ and $\beta\in\mathbb{R}^X$: $\alpha\geq\beta \Leftrightarrow \forall x\in X, \alpha(x) \geq \beta(x)$. We will often slightly abuse the notations and consider (where applicable) most objects as if they were matrices and vectors indexed by the set they operate upon.

We work with finite MDPs \cite{puterman1994markov}, that is tuples $\{S,A,P,R,\gamma\}$. The state space is noted $S$, $A$ is a finite action space, $R\in\mathbb{R}^{S\times A}$ is a reward-function, $\gamma\in (0,1)$ is a discount factor and $P\in \Delta_{S}^{S\times A}$ is the Markovian dynamics of the MDP. Thus, for each $(s,a)\in S\times A$, $P(.|s,a)$ is a distribution over $S$ and $P(s'|s,a)$ is the probability to reach $s'$ by choosing action $a$ in state $s$. At each time step $t$, the agent uses the information encoded in the state $s_t\in S$ in order to choose an action $a_t \in A$ according to a (deterministic\footnote{We restrict ourselves here to deterministic policies, but the loss of generality is minimal as there exists at least one optimal deterministic policy.}) policy $\pi\in A^S$. The agent then steps to a new state $s_{t+1}\in S$ according to the Markovian transition probabilities $P(s_{t+1}|s_t,a_t)$. Given that $P_\pi = (P(s'|s,\pi(s)))_{s,s' \in S}$ is the irreductible transition probability matrix, the stationary distribution over the states $\rho_\pi$ induced by a policy $\pi$ satisfies $\rho_\pi^TP_\pi = \rho_\pi^T\textrm{, with }X^T\textrm{ being the transpose of }X$.
The stationary distribution relative to the exert policy $\pi_E$ is $\rho_E$.

@@

When the state space is too large to allow matrix representations or when the transition probabilites or even the reward function are unknown except through observations gained by interacting with the system, RL or ADP may be used to approximate the optimal control policy~\cite{sutton1998reinforcement}.

We recall that solving the MDP is the direct problem. This contribution aims at solving the inverse one. We observe trajectories drawn from an expert's deterministic\footnotemark[\value{footnote}] policy $\pi_E$, assuming that there exists some unknown reward $R_E$ for which the expert is optimal. The suboptimality of the expert is an interesting setting that has been discussed for example in \cite{melo2010analysis,syed2010reduction}, but that we are not addressing here. We do not try to find this unknown reward $R_E$ but rather a non trivial reward $R$ for which the expert is at least near-optimal. The trivial reward $0$ is a solution to this ill-posed problem (no reward means that every behavior is optimal). Because of its ill-posed nature, this expression of \emph{Inverse Reinforcement Learning} (IRL) still has to find a satisfactory solution although a lot of progress has been made, see Sec.~\ref{sec:related}.
\section{The cascading algorithm}
\label{sec:algo}
Our first step towards a reward function solving the IRL problem is a classification step using a \emph{score function-based multi-class classifier} (SFMC$^2$ for short). This classifier learns a score function $q\in\mathbb{R}^{S\times A}$ that rates the association of a given action\footnote{Here, actions play the role of what is known as {\it labels} or {\it categories} when talking about classifiers.} $a\in A$ with a certain input $s\in S$. The classification rule $\pi_C\in A^S$ simply selects (one of) the action(s) that achieves the highest score for the given inputs:
\begin{equation}
  \label{eq:greedy2}
\pi_C(s) \in \argmax_a q(s,a).
\end{equation}
For example, \emph{Multi-class Support Vector Machines}~\cite{guermeur2011generic} can be seen as SFMC$^2$ algorithms, the same can be said of the structured margin approach \cite{taskar2005learning} both of which we consider in the experimental setting. Other algorithms may be envisionned (see Sec.~\ref{subsubsec:subroutines}).

Given a dataset $D_C = \{(s_i,a_i=\pi_E(s_i))_i\}$ of actions $a_i$ (deterministically) chosen by the expert on states $s_i$, we train such a classifier. The classification policy $\pi_C$ is not the end product we are looking for (that would be mere supervised imitation of the expert, not IRL). What is of particular interest to us is the score function $q$ itself. One can easily notice the similarity between Eq.~\eqref{eq:greedy2} and Eq.~\eqref{eq:greedy} that describes the relation between the optimal policy in an MDP and its optimal action-value function. The score function $q$ of the classifier can thus be viewed as some kind of optimal action-value function for the classifier policy $\pi_C$. By inversing the Bellman equation \eqref{eq:bellman1} with $q$ in lieu of $Q^\pi_R$, one gets $R^C$, the reward function relative to our score/action-value function $q$:
\begin{equation}
  \label{eq:rc}
  R^C(s,a) =q(s,a) - \sum_{\mathrm{s'}}P(s'|s,a)q(s',\pi_C(s')).
\end{equation}
As we wish to approximately solve the general IRL problem where the transition probabilities $P$ are unknown, our reward function $R^C$ will be approximated with the help of information gathered by interacting with the system. We assume that another dataset $D_R = \{(s_j,a_j,s'_j)_j\}$ is available where $s'_j$ is the state an agent taking action $a_j$ in state $s_j$ transitioned to. Action $a_j$ need not be chosen by any particular policy. The dataset $D_R$ brings us information about the dynamics of the system. From it, we construct datapoints
\begin{equation}
  \label{eq:rj}
  \{\hat r_j = q(s_j,a_j) - \gamma q(s'_j,\pi_C(s'_j))\}_j.
\end{equation}
As $s'_j$ is sampled accordingly to $P(\cdot|s_j,a_j)$ the constructed datapoints help building a good approximation of $R^C(s_j,a_j)$. A regressor (a simple least-square approximator can do but other solutions could also be envisionned, see Sec.~\ref{subsubsec:subroutines}) is then fed the datapoints $((s_j,a_i),\hat r_j)$ to obtain $\hat R^C$, a generalisation of $\{((s_j,a_j),\hat r_j)_j\}$ over the whole state-action space. The complete algorithm is given in Alg.~\ref{algo:cascading}.

@@

Cascading two supervised approaches like we do is a way to inject the MDP structure into the resolution of the problem. Indeed, mere classification only takes into account information from the expert (i.e., which action goes with which state) whereas using the Bellman equation in the expression of $\hat r_j$ makes use of the information lying in the transitions $(s_j,a_j,s'_j)$, namely information about the transition probabilities $P$. The final regression step is a way to generalize this information about $P$ to the whole state-action space in order to have a well-behaved reward function.

Being able to alleviate the ill effects of scalability or data scarcity by leveraging the wide range of techniques developped for the classification and regression problems is a strong advantage of the CSI approach.

@@

In this section, we prove that the deterministic expert policy $\pi_E$ is near optimal for the reward $\hat{R}^C$ the regression step outputs.
More formally, recalling from Sec.~\ref{sec:background} that $\rho_E$ is the stationary distribution of the expert policy, we prove that $\E_{s\sim\rho_E}[V^*_{\hat{R}^C}(s)-V^{\pi_E}_{\hat{R}^C}(s)]$ is bounded by a term that depends on:
\begin{itemize}
  \item the classification error defined as
$\epsilon_C=\E_{s\sim\rho_E}[\mathds{1}_{\{\pi_C(s)\neq\pi_E(s)\}}]$;
\item the regression error defined as 
 $   \epsilon_R=\max_{\pi\in A^S}\|\epsilon_\pi^R\|_{1,\rho_E}$,
  with:
  \begin{itemize}
  \item the subscript notation already used for $R_\pi$ and $P_\pi$ in Sec.~\ref{sec:background} meaning that, given an $X\in\mathbb{R}^{S\times A}$, $\pi\in A^S$, and $a\in A$,  $X_\pi\in\mathbb{R}^S$ and $X_a\in\mathbb{R}^S$ are respectively such that: $\forall s\in S, X_\pi(s)=X(s,\pi(s))$ and $\forall s\in S, X_a(s)=X(s,a)$ ;
  \item $\epsilon^R_\pi=R^C_\pi-\hat{R}^C_\pi$ ;
  \item $\|.\|_{1,\mu}$ the $\mu$-weighted $L_1$ norm: $\|f\|_{1,\mu} = \E_{x\sim \mu}[|f(x)|]$;
  \end{itemize}
\item the concentration coefficient $C_* = C_{\hat \pi_C}$ with:
  \begin{itemize}
  \item $C_{\pi}=(1-\gamma)\sum_{t\geq0}\gamma^tc_{\pi}(t), \text{ with } c_{\pi}(t)=\max_{s\in S}\frac{(\rho_E^TP^t_\pi)(s)}{\rho_E(s)}$ ;
  \item $\hat \pi_C$, the optimal policy for the reward $\hat R^C$ output by the algorithm ;
  \end{itemize}
  The constant $C_*$ can be estimated \emph{a posteriori} (after $\hat{R}^C$ is computed). {\it A priori}, $C_*$ can be upper-bounded by a more usual and general concentration coefficient but $C_*$ gives a tighter final result: one can informally see $C_*$ as a measure of the similarity between the distributions induced by $\hat\pi_C$ and $\pi_E$ (roughly, if $\hat{\pi}_C\approx\pi_E$ then $C_* \approx 1$).
\item $\Delta q = \max_{s\in S}(\max_{a\in A}q(s,a)-\min_{a\in A}q(s,a))=\max_{s\in S}(q(s,\pi_C(s))-\min_{a\in A}q(s,a))$,
    which could be normalized to $1$ without loss of generality. The range of variation of $q$ induces the one of $R^C$, $\hat{R}^C$ and $V^{\pi_C}_{\hat{R^C}}$.
  \end{itemize}
\begin{theorem}
\label{thm}
Let $\pi_E$ be the deterministic expert policy, $\rho_E$ its stationary distribution and $\hat{R}^C$ the reward the cascading algorithm outputs. We have:
\begin{equation}
0\leq\E_{s\sim\rho_E}[V^*_{\hat{R}^C}(s)-V^{\pi_E}_{\hat{R}^C}(s)]\leq \frac{1}{1-\gamma}\left(\epsilon_C\Delta q +\epsilon_R(1+C_*)\right).
\end{equation}
\end{theorem}
\begin{proof}
First let's recall some notations, $q\in\mathbb{R}^{S\times A}$ is the score function output by the classification step, $\pi_C$ is a deterministic classifier policy so that $\forall s \in S,\pi_C(s)\in\argmax_{a\in A}q(s,a)$, $R^C\in\mathbb{R}^{S\times A}$ is so that $\forall (s,a)\in S\times A, R^C(s,a)=q(s,a) -\gamma\sum_{s'\in S}P(s'|s,a)q(s',\pi_C(s'))$,
and $\hat{R}^C\in\mathbb{R}^{S\times A}$ is the reward function output by the regression step.

The difference between $R^C$ and $\hat{R}^C$ is noted $\epsilon^R=R^C-\hat{R}^C$.
We also introduce the reward function $R^E\in\mathbb{R}^{S\times A}$ which will be useful in our proof, not to be confused with $R_E$ the unknown reward function the expert optimizes:
\begin{equation}
\forall (s,a)\in S\times A, R^E(s,a)=q(s,a) -\gamma\sum_{s'\in S}P(s'|s,a)q(s',\pi_E(s')).
\end{equation}

We no have the following vectorial equalities $
R^C_a=q_a-\gamma P_aq_{\pi_C} ;
R^E_a=q_a-\gamma P_aq_{\pi_E};
\epsilon^R_a=R^C_a-\hat{R}^C_a$.
Now, we are going to upper bound the term: $\E_{s\sim\rho_E}[V^*_{\hat{R}^C}-V^{\pi_E}_{\hat{R}^C}]\geq0$ (the lower bound is obvious as $V^*$ is optimal).
Recall that $\hat{\pi}_C$ is a deterministic optimal policy of the reward $\hat{R}^C$. First, the term $V^*_{\hat{R}^C}-V^{\pi_E}_{\hat{R}^C}$ is decomposed:
\begin{equation}
V^*_{\hat{R}^C}-V^{\pi_E}_{\hat{R}^C}=(V^{\hat{\pi}_C}_{\hat{R}^C}-V^{\hat{\pi}_C}_{R^C})+(V^{\hat{\pi}_C}_{R^C}-V^{\pi_E}_{R^C})+(V^{\pi_E}_{R^C}-V^{\pi_E}_{\hat{R}^C}).
\end{equation}
We are going to bound each of these three terms. First, let $\pi$ be a given deterministic policy. We have, using $\epsilon^R = R^C - \hat R^C$:
$V^{\pi}_{R^C}-V^{\pi}_{\hat{R}^C}=V^{\pi}_{\epsilon^R}=(I-\gamma P_\pi)^{-1}\epsilon^R_{\pi}$.
If $\pi=\pi_E$, we have, thanks to the power series expression of $(I-\gamma P_{\pi_E})^{-1}$, the definition of $\rho_E$ and the definition of the $\mu$-weighted $L_1$ norm, one property of which is that $\forall X, \mu^TX \leq \|X\|_{1,\mu}$:
\begin{equation}
  \rho_E^T(V^{\pi}_{R^C}-V^{\pi}_{\hat{R}^C})=\rho_E^T(I-\gamma P_{\pi_E})^{-1}\epsilon^R_{\pi_E}=\frac{1}{1-\gamma}\rho_E^T\epsilon^R_{\pi_E}\leq\frac{1}{1-\gamma}\|\epsilon^R_{\pi_E}\|_{1,\rho_E}.\end{equation}

@@

So, using the notations introduced before we stated the theorem, we are able to give an upper bound to the first and third terms (recall also the notation $C_{\hat \pi_C} = C_*$):

@@

V^{\hat{\pi}_C}_{R^C}-V^{\pi_E}_{R^C}=(V^{\hat{\pi}_C}_{R^C}-V^{\pi_C}_{R^C})+(V^{\pi_C}_{R^C}-V^{\pi_E}_{R^E})+(V^{\pi_E}_{R^E}-V^{\pi_E}_{R^C}).

@@

V^{\hat{\pi}_C}_{R^C}-V^{\pi_E}_{R^C}\leq(V^{\pi_C}_{R^C}-V^{\pi_E}_{R^E})+(V^{\pi_E}_{R^E}-V^{\pi_E}_{R^C}).

@@

By construction, we have $V^{\pi_C}_{R^C}=q_{\pi_C}$ and $V^{\pi_E}_{R^E}=q_{\pi_E}$, thus:

@@

\rho_E^T(V^{\pi_C}_{R^C}-V^{\pi_E}_{R^E})&=&\rho_E^T(q_{\pi_C}-q_{\pi_E})\\

@@

$\rho_E^T(V^{\pi_C}_{R^C}-V^{\pi_E}_{R^E})\leq\Delta q\sum_{s\in S}\rho_E(s)[\mathds{1}_{\{\pi_C(s)\neq\pi_E(s)\}}]= \Delta q \epsilon_C$.

@@

\rho_E^T(V^{\pi_E}_{R^E}-V^{\pi_E}_{R^C})&=\rho_E^T(I-\gamma P_{\pi_E})^{-1}(R^E_{\pi_E}-R^C_{\pi_E})=\rho_E^T(I-\gamma P_{\pi_E})^{-1}\gamma P_{\pi_E}(q_{\pi_C}-q_{\pi_E}),

@@

%\qed

@@

The attentive reader, however, will percieve that similarities stop there. The error terms occurring in the two bounds are not related to one another. As CSI makes no use of the feature expectation of the expert, what is known as $\bar\epsilon_Q$ in \cite{klein2012scirl} does not appear in this analysis. Likewise, the regression error $\epsilon_R$ of this paper does not appear in the analysis of SCIRL, which does not use a regressor. Perhaps more subtly, the classification error and classification policy known in both papers as $\epsilon_C$ and $\pi_C$ are not the same. The classification policy of SCIRL is no tantamount to what is called $\pi_C$ here. For SCIRL, $\pi_C$ is the greedy policy for an approximation of the value function of the expert with respect to the reward output by the algorithm. For CSI, $\pi_C$ is the decision rule of the classifier, an object that is not aware of the structure of the MDP. We shall also mention that the error terms appearing in the CSI bound are more standard than the ones of SCIRL ({\it e.g.}, regression error vs feature expectation estimation error) thus they may be easier to control.

@@

Furthermore, the classifier whose results are diplayed here is the output of the first step of the algorithm. The classification performance is obviously not that good, this hints to the fact that our algorithm may be empirically more forgiving of classification errors than our theoretical bound lets us expect.

>>>>

IRL is one of many ways to perform Apprenticeship Learning (AL): imitating a demonstrated expert policy, without necessarily explicitly looking for the reward function. The reward function nevertheless is of interest in its own right. As mentioned in \cite{russell1998learning}, its semantics can be analyzed in biology or econometrics for instance. Practically, the reward can be seen as a succinct description of a task. Discovering it removes the coupling that exists in AL between understanding the task and learning how to fulfill it. IRL allows the use of (A)DP or RL techniques to learn how to do the task from the computed reward function. A very straightforward non-IRL way to do AL is for example to use a multi-class classifier to directly learn the expert policy. We provide in the experiments (Sec.~\ref{sec:experiments}) a comparison between AL and IRL algorithms by using IRL as a way to do AL.

@@

This algorithm does not need to iteratively solve an MDP and requires only sampled transitions from expert and non-expert policies as inputs. Moreover, up to the use of some heuristics (see Sec.~\ref{subsubsec:heuristics}), the algorithm is able to be trained only with transitions sampled from the demonstrated expert policy. A specific instantiation of CSI (proposed in Sec.~\ref{subsubsec:subroutines}) is tested on the mountain car problem (Sec.~\ref{subsec:mountaincar}) and on a highway driving simulator (Sec.~\ref{subsec:highway}) where we compare it with a pure classification algorithm \cite{taskar2005learning} and with two recent successful IRL methods \cite{klein2012scirl} as well as with a random baseline. Differences and similarities with existing AL or IRL approaches are succinctly discussed in Sec.~\ref{sec:related}.
\section{Background and notation}
\label{sec:background}
First, we introduce some general notation.
Let $E$ and $F$ be two non-empty sets, $E^F$ is the set of functions from $F$ to $E$.
We note $\Delta_X$ the set of distributions over $X$.
Let $\alpha\in\mathbb{R}^X$ and $\beta\in\mathbb{R}^X$: $\alpha\geq\beta \Leftrightarrow \forall x\in X, \alpha(x) \geq \beta(x)$. We will often slightly abuse the notation and consider (where applicable) most objects as if they were matrices and vectors indexed by the set they operate upon.

We work with finite MDPs \cite{puterman1994markov}, that is tuples $\{S,A,P,R,\gamma\}$. The state space is noted $S$, $A$ is a finite action space, $R\in\mathbb{R}^{S\times A}$ is a reward-function, $\gamma\in (0,1)$ is a discount factor and $P\in \Delta_{S}^{S\times A}$ is the Markovian dynamics of the MDP. Thus, for each $(s,a)\in S\times A$, $P(.|s,a)$ is a distribution over $S$ and $P(s'|s,a)$ is the probability to reach $s'$ by choosing action $a$ in state $s$. At each time step $t$, the agent uses the information encoded in the state $s_t\in S$ in order to choose an action $a_t \in A$ according to a (deterministic\footnote{We restrict ourselves here to deterministic policies, but the loss of generality is minimal as there exists at least one optimal deterministic policy.}) policy $\pi\in A^S$. The agent then steps to a new state $s_{t+1}\in S$ according to the Markovian transition probabilities $P(s_{t+1}|s_t,a_t)$. Given that $P_\pi = (P(s'|s,\pi(s)))_{s,s' \in S}$ is the transition probability matrix, the stationary distribution over the states $\rho_\pi$ induced by a policy $\pi$ satisfies $\rho_\pi^TP_\pi = \rho_\pi^T\textrm{, with }X^T\textrm{ being the transpose of }X$.
The stationary distribution relative to the expert policy $\pi_E$ is $\rho_E$.

@@

When the state space is too large to allow matrix representations or when the transition probabilities or even the reward function are unknown except through observations gained by interacting with the system, RL or ADP may be used to approximate the optimal control policy~\cite{sutton1998reinforcement}.

We recall that solving the MDP is the direct problem. This contribution aims at solving the inverse one. We observe trajectories drawn from an expert's deterministic\footnotemark[\value{footnote}] policy $\pi_E$, assuming that there exists some unknown reward $R_E$ for which the expert is optimal. The suboptimality of the expert is an interesting setting that has been discussed for example in \cite{melo2010analysis,syed2010reduction}, but that we are not addressing here. We do not try to find this unknown reward $R_E$ but rather a non trivial reward $R$ for which the expert is at least near-optimal. The trivial reward $0$ is a solution to this ill-posed problem (no reward means that every behavior is optimal). Because of its ill-posed nature, this expression of \emph{Inverse Reinforcement Learning} (IRL) still has to find a satisfactory solution although a lot of progress has been made, see Sec.~\ref{sec:related}.
\section{The cascading algorithm}
\label{sec:algo}
Our first step towards a reward function solving the IRL problem is a classification step using a \emph{score function-based multi-class classifier} (SFMC$^2$ for short). This classifier learns a score function $q\in\mathbb{R}^{S\times A}$ that rates the association of a given action\footnote{Here, actions play the role of what is known as {\it labels} or {\it categories} when talking about classifiers.} $a\in A$ with a certain input $s\in S$. The classification rule $\pi_C\in A^S$ simply selects (one of) the action(s) that achieves the highest score for the given inputs:
\begin{equation}
  \label{eq:greedy2}
\pi_C(s) \in \argmax_a q(s,a).
\end{equation}
For example, \emph{Multi-class Support Vector Machines}~\cite{guermeur2011generic} can be seen as SFMC$^2$ algorithms, the same can be said of the structured margin approach \cite{taskar2005learning} both of which we consider in the experimental setting. Other algorithms may be envisioned (see Sec.~\ref{subsubsec:subroutines}).

Given a dataset $D_C = \{(s_i,a_i=\pi_E(s_i))_i\}$ of actions $a_i$ (deterministically) chosen by the expert on states $s_i$, we train such a classifier. The classification policy $\pi_C$ is not the end product we are looking for (that would be mere supervised imitation of the expert, not IRL). What is of particular interest to us is the score function $q$ itself. One can easily notice the similarity between Eq.~\eqref{eq:greedy2} and Eq.~\eqref{eq:greedy} that describes the relation between the optimal policy in an MDP and its optimal action-value function. The score function $q$ of the classifier can thus be viewed as some kind of optimal action-value function for the classifier policy $\pi_C$. By inversing the Bellman equation \eqref{eq:bellman1} with $q$ in lieu of $Q^\pi_R$, one gets $R^C$, the reward function relative to our score/action-value function $q$:
\begin{equation}
  \label{eq:rc}
  R^C(s,a) =q(s,a) -\gamma \sum_{\mathrm{s'}}P(s'|s,a)q(s',\pi_C(s')).
\end{equation}
As we wish to approximately solve the general IRL problem where the transition probabilities $P$ are unknown, our reward function $R^C$ will be approximated with the help of information gathered by interacting with the system. We assume that another dataset $D_R = \{(s_j,a_j,s'_j)_j\}$ is available where $s'_j$ is the state an agent taking action $a_j$ in state $s_j$ transitioned to. Action $a_j$ need not be chosen by any particular policy. The dataset $D_R$ brings us information about the dynamics of the system. From it, we construct datapoints
\begin{equation}
  \label{eq:rj}
  \{\hat r_j = q(s_j,a_j) - \gamma q(s'_j,\pi_C(s'_j))\}_j.
\end{equation}
As $s'_j$ is sampled according to $P(\cdot|s_j,a_j)$ the constructed datapoints help building a good approximation of $R^C(s_j,a_j)$. A regressor (a simple least-square approximator can do but other solutions could also be envisioned, see Sec.~\ref{subsubsec:subroutines}) is then fed the datapoints $((s_j,a_i),\hat r_j)$ to obtain $\hat R^C$, a generalization of $\{((s_j,a_j),\hat r_j)_j\}$ over the whole state-action space. The complete algorithm is given in Alg.~\ref{algo:cascading}.

@@

Cascading two supervised approaches like we do is a way to inject the MDP structure into the resolution of the problem. Indeed, mere classification only takes into account information from the expert (i.e., which action goes with which state) whereas using the Bellman equation in the expression of $\hat r_j$ makes use of the information lying in the transitions $(s_j,a_j,s'_j)$, namely information about the transition probabilities $P$. The final regression step is a way to generalize this information about $P$ to the whole state-action space in order to have a well-behaved reward function. Being able to alleviate the ill effects of scalability or data scarcity by leveraging the wide range of techniques developed for the classification and regression problems is a strong advantage of the CSI approach.

@@

In this section, we prove that the deterministic expert policy $\pi_E$ is near optimal for the reward $\hat{R}^C$ the regression step outputs.
More formally, recalling from Sec.~\ref{sec:background} that $\rho_E$ is the stationary distribution of the expert policy, we prove that $\E_{s\sim\rho_E}[V^*_{\hat{R}^C}(s)-V^{\pi_E}_{\hat{R}^C}(s)]$ is bounded by a term that depends on:
\begin{itemize}
  \item the classification error defined as
$\epsilon_C=\E_{s\sim\rho_E}[\mathds{1}_{\{\pi_C(s)\neq\pi_E(s)\}}]$;
\item the regression error defined as 
 $   \epsilon_R=\max_{\pi\in A^S}\|\epsilon_\pi^R\|_{1,\rho_E}$,
  with:
  \begin{itemize}
  \item the subscript notation already used for $R_\pi$ and $P_\pi$ in Sec.~\ref{sec:background} meaning that, given an $X\in\mathbb{R}^{S\times A}$, $\pi\in A^S$, and $a\in A$,  $X_\pi\in\mathbb{R}^S$ and $X_a\in\mathbb{R}^S$ are respectively such that: $\forall s\in S, X_\pi(s)=X(s,\pi(s))$ and $\forall s\in S, X_a(s)=X(s,a)$ ;
  \item $\epsilon^R_\pi=R^C_\pi-\hat{R}^C_\pi$ ;
  \item $\|.\|_{1,\mu}$ the $\mu$-weighted $L_1$ norm: $\|f\|_{1,\mu} = \E_{x\sim \mu}[|f(x)|]$;
  \end{itemize}
\item the concentration coefficient $C_* = C_{\hat \pi_C}$ with:
  \begin{itemize}
  \item $C_{\pi}=(1-\gamma)\sum_{t\geq0}\gamma^tc_{\pi}(t), \text{ with } c_{\pi}(t)=\max_{s\in S}\frac{(\rho_E^TP^t_\pi)(s)}{\rho_E(s)}$ ;
  \item $\hat \pi_C$, the optimal policy for the reward $\hat R^C$ output by the algorithm ;
  \end{itemize}
  The constant $C_*$ can be estimated \emph{a posteriori} (after $\hat{R}^C$ is computed). {\it A priori}, $C_*$ can be upper-bounded by a more usual and general concentration coefficient but $C_*$ gives a tighter final result: one can informally see $C_*$ as a measure of the similarity between the distributions induced by $\hat\pi_C$ and $\pi_E$ (roughly, if $\hat{\pi}_C\approx\pi_E$ then $C_* \approx 1$).
\item $\Delta q = \max_{s\in S}(\max_{a\in A}q(s,a)-\min_{a\in A}q(s,a))=\max_{s\in S}(q(s,\pi_C(s))-\min_{a\in A}q(s,a))$,
    which could be normalized to $1$ without loss of generality. The range of variation of $q$ is tied to the one of $R^C$, $\hat{R}^C$ and $V^{\pi_C}_{\hat{R^C}}$. What matters with these objects is the relative values for different state action couples, not the objective range. They can be shifted and positively scaled without consequence.
  \end{itemize}
\begin{theorem}
\label{thm}
Let $\pi_E$ be the deterministic expert policy, $\rho_E$ its stationary distribution and $\hat{R}^C$ the reward the cascading algorithm outputs. We have:
\begin{equation}
0\leq\E_{s\sim\rho_E}[V^*_{\hat{R}^C}(s)-V^{\pi_E}_{\hat{R}^C}(s)]\leq \frac{1}{1-\gamma}\left(\epsilon_C\Delta q +\epsilon_R(1+C_*)\right).
\end{equation}
\end{theorem}
\begin{proof}
First let's recall some notation, $q\in\mathbb{R}^{S\times A}$ is the score function output by the classification step, $\pi_C$ is a deterministic classifier policy so that $\forall s \in S,\pi_C(s)\in\argmax_{a\in A}q(s,a)$, $R^C\in\mathbb{R}^{S\times A}$ is so that $\forall (s,a)\in S\times A, R^C(s,a)=q(s,a) -\gamma\sum_{s'\in S}P(s'|s,a)q(s',\pi_C(s'))$,
and $\hat{R}^C\in\mathbb{R}^{S\times A}$ is the reward function output by the regression step.

The difference between $R^C$ and $\hat{R}^C$ is noted $\epsilon^R=R^C-\hat{R}^C\in\mathbb{R}^{S\times A}$.
We also introduce the reward function $\mathfrak{R}^E\in\mathbb{R}^{S\times A}$ which will be useful in our proof, not to be confused with $R_E$ the unknown reward function the expert optimizes:
\begin{equation}
\forall (s,a)\in S\times A, \mathfrak{R}^E(s,a)=q(s,a) -\gamma\sum_{s'\in S}P(s'|s,a)q(s',\pi_E(s')).
\end{equation}

We now have the following vectorial equalities $
R^C_a=q_a-\gamma P_aq_{\pi_C} ;
\mathfrak{R}^E_a=q_a-\gamma P_aq_{\pi_E};
\epsilon^R_a=R^C_a-\hat{R}^C_a$.
Now, we are going to upper bound the term: $\E_{s\sim\rho_E}[V^*_{\hat{R}^C}-V^{\pi_E}_{\hat{R}^C}]\geq0$ (the lower bound is obvious as $V^*$ is optimal).
Recall that $\hat{\pi}_C$ is a deterministic optimal policy of the reward $\hat{R}^C$. First, the term $V^*_{\hat{R}^C}-V^{\pi_E}_{\hat{R}^C}$ is decomposed:
\begin{equation}
V^*_{\hat{R}^C}-V^{\pi_E}_{\hat{R}^C}=(V^{\hat{\pi}_C}_{\hat{R}^C}-V^{\hat{\pi}_C}_{R^C})+(V^{\hat{\pi}_C}_{R^C}-V^{\pi_E}_{R^C})+(V^{\pi_E}_{R^C}-V^{\pi_E}_{\hat{R}^C}).
\end{equation}
We are going to bound each of these three terms. First, let $\pi$ be a given deterministic policy. We have, using $\epsilon^R = R^C - \hat R^C$:
$V^{\pi}_{R^C}-V^{\pi}_{\hat{R}^C}=V^{\pi}_{\epsilon^R}=(I-\gamma P_\pi)^{-1}\epsilon^R_{\pi}$.
If $\pi=\pi_E$, we have, thanks to the power series expression of $(I-\gamma P_{\pi_E})^{-1}$, the definition of $\rho_E$ and the definition of the $\mu$-weighted $L_1$ norm, one property of which is that $\forall X, \mu^TX \leq \|X\|_{1,\mu}$:
\begin{equation}
  \rho_E^T(V^{\pi}_{R^C}-V^{\pi}_{\hat{R}^C})=\rho_E^T(I-\gamma P_{\pi_E})^{-1}\epsilon^R_{\pi_E}=\frac{1}{1-\gamma}\rho_E^T\epsilon^R_{\pi_E}\leq\frac{1}{1-\gamma}\|\epsilon^R_{\pi_E}\|_{1,\rho_E}.\end{equation}

@@

So, using the notation introduced before we stated the theorem, we are able to give an upper bound to the first and third terms (recall also the notation $C_{\hat \pi_C} = C_*$):

@@

V^{\hat{\pi}_C}_{R^C}-V^{\pi_E}_{R^C}=(V^{\hat{\pi}_C}_{R^C}-V^{\pi_C}_{R^C})+(V^{\pi_C}_{R^C}-V^{\pi_E}_{\mathfrak{R}^E})+(V^{\pi_E}_{\mathfrak{R}^E}-V^{\pi_E}_{R^C}).

@@

V^{\hat{\pi}_C}_{R^C}-V^{\pi_E}_{R^C}\leq(V^{\pi_C}_{R^C}-V^{\pi_E}_{\mathfrak{R}^E})+(V^{\pi_E}_{\mathfrak{R}^E}-V^{\pi_E}_{R^C}).

@@

By construction, we have $V^{\pi_C}_{R^C}=q_{\pi_C}$ and $V^{\pi_E}_{\mathfrak{R}^E}=q_{\pi_E}$, thus:

@@

\rho_E^T(V^{\pi_C}_{R^C}-V^{\pi_E}_{\mathfrak{R}^E})&=&\rho_E^T(q_{\pi_C}-q_{\pi_E})\\

@@

$\rho_E^T(V^{\pi_C}_{R^C}-V^{\pi_E}_{\mathfrak{R}^E})\leq\Delta q\sum_{s\in S}\rho_E(s)[\mathds{1}_{\{\pi_C(s)\neq\pi_E(s)\}}]= \Delta q \epsilon_C$.

@@

\rho_E^T(V^{\pi_E}_{\mathfrak{R}^E}-V^{\pi_E}_{R^C})&=\rho_E^T(I-\gamma P_{\pi_E})^{-1}(\mathfrak{R}^E_{\pi_E}-R^C_{\pi_E})=\rho_E^T(I-\gamma P_{\pi_E})^{-1}\gamma P_{\pi_E}(q_{\pi_C}-q_{\pi_E}),

@@

\qed

@@

The attentive reader, however, will perceive that similarities stop there. The error terms occurring in the two bounds are not related to one another. As CSI makes no use of the feature expectation of the expert, what is known as $\bar\epsilon_Q$ in \cite{klein2012scirl} does not appear in this analysis. Likewise, the regression error $\epsilon_R$ of this paper does not appear in the analysis of SCIRL, which does not use a regressor. Perhaps more subtly, the classification error and classification policy known in both papers as $\epsilon_C$ and $\pi_C$ are not the same. The classification policy of SCIRL is not tantamount to what is called $\pi_C$ here. For SCIRL, $\pi_C$ is the greedy policy for an approximation of the value function of the expert with respect to the reward output by the algorithm. For CSI, $\pi_C$ is the decision rule of the classifier, an object that is not aware of the structure of the MDP. We shall also mention that the error terms appearing in the CSI bound are more standard than the ones of SCIRL ({\it e.g.}, regression error vs feature expectation estimation error) thus they may be easier to control.

@@

Furthermore, the classifier whose results are displayed here is the output of the first step of the algorithm. The classification performance is obviously not that good, which points to the fact that our algorithm may be empirically more forgiving of classification errors than our theoretical bound lets us expect.
